### 2022-11-24 dissertation/chapters/01-introduction.tex
✨ (final-review): Up to chapter 4

\begin{quotation}
      \noindent ``It is better to solve one problem five different ways, than to solve five problems one way.''
\end{quotation}
\begin{flushright}
      - George Polya
\end{flushright}

@@

This chapter provides the reader with a brief overview of the problem domain and outlines the research problem, objectives and purpose. The remainder of the chapter is presented as follows.

@@

The remainder of this dissertation is structured as follows.

>>>>

\begin{quotation}
      ``It is better to solve one problem five different ways, than to solve five problems one way.''
\end{quotation}
\begin{flushright}
      - George Polya
\end{flushright}

@@

This chapter provides the reader with a brief overview of the problem domain and outlines the research problem, objectives and purpose. The remainder of the chapter is presented as follows:

@@

The remainder of this dissertation is structured as follows:

### 2022-11-24 dissertation/chapters/02-artificial-neural-networks.tex
✨ (final-review): Up to chapter 4

The human brain is one of the most complicated biological organs in nature. It gives us the ability to think and learn. The field of \acs{ML} has taken a lot of inspiration from the biological brain which lead to the development of \acp{ANN}~\cite{ref:rosenblatt:1958}. \acp{ANN} are used throughout this dissertation as the \textit{model} to be trained using \acp{HH}. The purpose of this chapter is to provide the necessary background information needed on \acp{ANN} and is structured as follows.

@@

\acp{ANN} are inspired by the biological brain. Our brains model biological neural networks with neuron counts in the hundreds of billions and synapse counts in the trillions. To emulate the computational capability of the human brain is not yet possible on modern day hardware.~\citeauthor{ref:sandberg:2008}~\cite{ref:sandberg:2008} approximates a computational requirement of 256 000 terabytes/s to emulate the entire brain. Despite our shortage in hardware capabilities, \acp{ANN} have been successfully applied to a range of problem classes.~\citeauthor{ref:engelbrecht:2007}~\cite{ref:engelbrecht:2007} summarises some common problems that are solved using \acp{ANN} and can be listed as follows.

>>>>

The human brain is one of the most complicated biological organs in nature. It gives us the ability to think and learn. The field of \acs{ML} has taken a lot of inspiration from the biological brain which lead to the development of \acp{ANN}~\cite{ref:rosenblatt:1958}. \acp{ANN} are used throughout this dissertation as the \textit{model} to be trained using \acp{HH}. The purpose of this chapter is to provide the necessary background information needed on \acp{ANN} and is structured as follows:

@@

\acp{ANN} are inspired by the biological brain. Our brains model biological neural networks with neuron counts in the hundreds of billions and synapse counts in the trillions. To emulate the computational capability of the human brain is not yet possible on modern day hardware.~\citeauthor{ref:sandberg:2008}~\cite{ref:sandberg:2008} approximates a computational requirement of 256 000 terabytes/s to emulate the entire brain. Despite our shortage in hardware capabilities, \acp{ANN} have been successfully applied to a range of problem classes.~\citeauthor{ref:engelbrecht:2007}~\cite{ref:engelbrecht:2007} summarises some common problems that are solved using \acp{ANN} and can be listed as follows:

### 2022-11-24 dissertation/chapters/03-heuristics.tex
✨ (final-review): Up to chapter 4

Chapter~\ref{chap:anns} briefly introduced the reader to the concept of \index{heuristic}\index{heuristic}heuristics and \acfp{MH}s. This chapter presents more detailed background information on various different \index{heuristic}heuristics that have been used to train \acp{FFNN}. Broadly speaking, this chapter focuses on two different groups of \index{heuristic}heuristics, including classical gradient-based approaches and population-based \acp{MH}. Each technique is presented and discussed in detail. Pseudo-code algorithms are provided for each technique and discussions follow on advantages, disadvantages, capabilities and limitations. The remainder of this chapter is structured as follows.

@@

The purpose of the \index{mutation}mutation operator is to produce a trial vector for each entity in the current population by mutating a \textit{target} vector with a \textit{weighted differential}~\cite{ref:engelbrecht:2007}. The trial vector is used in the \index{crossover}crossover operator to produce offspring. The \index{mutation}mutation process then follows as such. For each each parent $\boldsymbol{x}_{i}(t)$ generate a trial vector $\boldsymbol{u}_{i}(t)$ as follows.

>>>>

Chapter~\ref{chap:anns} briefly introduced the reader to the concept of \index{heuristic}\index{heuristic}heuristics and \acfp{MH}s. This chapter presents more detailed background information on various different \index{heuristic}heuristics that have been used to train \acp{FFNN}. Broadly speaking, this chapter focuses on two different groups of \index{heuristic}heuristics, including classical gradient-based approaches and population-based \acp{MH}. Each technique is presented and discussed in detail. Pseudo-code algorithms are provided for each technique and discussions follow on advantages, disadvantages, capabilities and limitations. The remainder of this chapter is structured as follows:

@@

The purpose of the \index{mutation}mutation operator is to produce a trial vector for each entity in the current population by mutating a \textit{target} vector with a \textit{weighted differential}~\cite{ref:engelbrecht:2007}. The trial vector is used in the \index{crossover}crossover operator to produce offspring. The \index{mutation}mutation process then follows as such. For each each parent $\boldsymbol{x}_{i}(t)$ generate a trial vector $\boldsymbol{u}_{i}(t)$ as follows:

### 2022-11-24 dissertation/chapters/04-hyper-heuristics.tex
✨ (final-review): Up to chapter 4

\begin{quotation}
      \noindent ``The capacity to learn is a gift; the ability to learn is a skill; the willingness to learn is a choice.''
\end{quotation}
\begin{flushright}
      - Brian Herbert
\end{flushright}

\noindent
Chapter \ref{chap:heuristics} introduced the concept of a \index{heuristic}heuristic as an optimisation algorithm that is used in a search process that seeks out optimal solutions to an optimisation problem. Finding the best \index{heuristic}heuristic to use for a given optimisation problem is non-trivial. Furthermore, finding optimal solutions is often dependent on using the appropriate configurations. In the context of training \acp{FFNN}, these configurations could include \index{heuristic}heuristic hyper-parameters and model architecture. It is often the case that the application of a specific configuration only applies to a specific problem and does not necessarily translate or generalise to other problems~\cite{ref:kheiri:2017}.

@@

More modern techniques automate the search for optimal configuration as part of the underlying search process itself, resulting in a form of \index{online learning}online learning. An example of such a technique is to dynamically adjust and learn the heuristic \index{hyper-parameters}\textit{hyper-parameters} as part of the optimisation process. This field of study is known as \index{meta-learning}meta-learning~\cite{ref:giraud:2004}.

A recent suggestion related to the field of \index{meta-learning}meta-learning is to dynamically select and/or adjust the \index{heuristic}heuristic used throughout the training process and not just the \index{heuristic}heuristic hyper-parameters. This approach focuses on the hybridisation of learning paradigms. By dynamically combining the best of different learning paradigms throughout the learning process, a trade-off can be made between focussing on a particular solution (exploitation), in comparison to seeking out novel solutions (exploration) during training. The dynamic trade-off between exploration and exploitation could also be seen as a dynamic trade-off between the strengths and weaknesses of different heuristics during the search process.

One such form of hybridisation of learning paradigms is referred to as \textit{heterogeneous learning approaches}. Heterogeneous learning approaches make use of different \textit{search behaviours} by selecting from a behaviour pool. Heterogeneous approaches have shown to balance the trade-off between exploration and exploitation~\cite{ref:nepomuceno:2013}.

Another form of hybridisation of learning paradigms is that of hybridisation of different \textit{heuristics} as they are applied to some optimisation problem~\cite{ref:burke:2013}. These methods are referred to as \acfp{HH} and focus on finding the best heuristic in \textit{heuristic space} to solve a specific problem. This dissertation takes a particular interest in developing a novel selection \acs{HH} that can be used to train \acp{FFNN}. This chapter provides the reader with the necessary background information of \index{meta-learning}meta-learning and \acp{HH}. The remainder of the chapter is structured as follows.

@@

\citeauthor{ref:vilalta:2002}~\cite{ref:vilalta:2002} mentions that meta-learning differs from base-learning in the scope of the level of adaptation. Meta-learning is concerned with learning how to choose the right configuration and biases dynamically. On the contrary, for \textit{base}-learning, biases and configurations are predefined and fixed a priori. \citeauthor{ref:thrun:1998}~\cite{ref:thrun:1998} mentions that meta-learning aims at discovering ways to dynamically search for the best learning strategy as the number of tasks increases, implying some form of generalisation of the learning process to other problems.

@@

\acp{HH} implement a form of \index{meta-learning}meta-learning that is concerned with the selection of the best heuristic from a pool of heuristics, to solve a given problem. In the context of population-based \acp{HH}, an entity pool exists that represent a pool of candidate solutions to the given problem. Each entity in the entity pool is assigned its own low-level \index{heuristic}heuristic from the heuristic pool.

@@

\noindent
Figure \ref{fig:heuristics:hh:hh_algorithm} shows the separation of information that is accessible to the entity pool and heuristic pool. At a higher level abstraction, the high level heuristic implemented by \acp{HH}, do not make use of domain specific information such as an entity's position or gradient from the search space.

@@

A \acs{HH} is considered to be a learning algorithm when it uses feedback from the search process to adapt and guide the search process. Learning \acp{HH} can implement a form of \index{online learning}\textit{online learning} or \index{offline learning}\textit{offline learning}.

@@

\acp{HH} that implement \index{offline learning}offline learning, implement a form of learning where knowledge, in the form of rules or programs, is gathered from a set of training instances and is then applied independently from the search process itself. Examples of offline learning approaches within hyper-heuristics include \textit{learning classifier systems}, \textit{case-based reasoning} and \acf{GP}.

@@

Selection \acp{HH} consists of two components, including a low-level \index{heuristic}heuristic selection strategy and a move acceptance strategy. Low-level \index{heuristic}heuristic selection can be done in a simple, \textit{non-adaptive} way. No learning is involved in these approaches. For non-adaptive techniques, heuristic selection is based on a predefined heuristic ordering that is generated either randomly or in a ordered cycle that is repeated throughout the optimisation process~\cite{ref:cowling:2000}. As an alternative, heuristic selection may incorporate an \textit{adaptive} (\index{online learning}\textit{online learning}) mechanism, based on the probabilistic weighting of the low-level heuristics~\cite{ref:burke:2003} or some type of performance statistics~\cite{ref:cowling:2000}.

A move acceptance strategy is the mechanism by which the application of a low-level \index{heuristic}heuristic is either accepted or rejected. In general, a move is accepted or rejected based on the quality of the move and the current solution during a single point search.~\citeauthor{ref:burke:2003}~\cite{ref:burke:2003} mentions that many move acceptance strategies have been explored within \acp{HH}.

@@

This chapter provided the reader with the necessary background information on \index{meta-learning}meta-learning and \acp{HH}. Formal definitions were given, followed by detailed discussions. A modern classification scheme for \acs{HH} as proposed by ~\citeauthor{ref:burke:2010}~\cite{ref:burke:2010} was presented in detail. Distinction is made between the source of feedback information used during learning and the nature of the heuristic search space. Detailed discussions on selection and generation heuristics are provided, followed by a further classification of construction and perturbation mechanisms.

>>>>

\begin{quotation}
      ``The capacity to learn is a gift; the ability to learn is a skill; the willingness to learn is a choice.''
\end{quotation}
\begin{flushright}
      - Brian Herbert
\end{flushright}

Chapter \ref{chap:heuristics} introduced the concept of a \index{heuristic}heuristic as an optimisation algorithm that is used in a search process that seeks out optimal solutions to an optimisation problem. Finding the best \index{heuristic}heuristic to use for a given optimisation problem is non-trivial. Furthermore, finding optimal solutions is often dependent on using the appropriate configurations. In the context of training \acp{FFNN}, these configurations could include \index{heuristic}heuristic hyper-parameters and model architecture. It is often the case that the application of a specific configuration only applies to a specific problem and does not necessarily translate or generalise to other problems~\cite{ref:kheiri:2017}.

@@

More modern techniques automate the search for optimal configuration as part of the underlying search process itself, resulting in a form of dynamic configuration during the search process. An example of such a technique is to dynamically adjust the heuristic \index{hyper-parameters}\textit{hyper-parameters} as part of the optimisation process. This field of study is known as \index{meta-learning}meta-learning~\cite{ref:giraud:2004}. The term \textit{learning} refers to the optimisation of the configuration and \index{hyper-parameters}hyper-parameters.

A recent suggestion related to the field of \index{meta-learning}meta-learning is to dynamically select and/or adjust the \index{heuristic}heuristic characteristics and heaviours used throughout the search process and not just the \index{heuristic}heuristic hyper-parameters. This approach focuses on the hybridisation of \index{heuristic}heuristic paradigms. By dynamically combining the best of different \index{heuristic} paradigms throughout the learning process, a trade-off can be made between focussing on a particular solution (exploitation), in comparison to seeking out novel solutions (exploration) during the search process. The dynamic trade-off between exploration and exploitation could also be seen as a dynamic trade-off between the strengths and weaknesses of different \index{heuristic}heuristics during the search process.

One such form of hybridisation of \index{heuristic}heuristic paradigms is referred to as \textit{heterogeneous approaches}. Heterogeneous approaches make use of different \textit{search behaviours} by selecting from a behaviour pool. Heterogeneous approaches have shown to balance the trade-off between exploration and exploitation~\cite{ref:nepomuceno:2013}.

Another form of hybridisation of \index{heuristic}heuristic paradigms is that of hybridisation of different \textit{heuristics} that are applied to some optimisation problem~\cite{ref:burke:2013}. These methods are referred to as \acfp{HH} and focus on finding the best heuristic in \textit{heuristic space} to solve a specific problem. This dissertation takes a particular interest in developing a novel selection \acs{HH} that can be used to train \acp{FFNN}. This chapter provides the reader with the necessary background information of \index{meta-learning}meta-learning and \acp{HH}. The remainder of the chapter is structured as follows:

@@

\citeauthor{ref:vilalta:2002}~\cite{ref:vilalta:2002} mention that meta-learning differs from base-learning in the scope of the level of adaptation. Meta-learning is concerned with learning how to choose the right configuration and biases dynamically. On the contrary, for \textit{base}-learning, biases and configurations are predefined and fixed a priori. \citeauthor{ref:thrun:1998}~\cite{ref:thrun:1998} mention that meta-learning aims at discovering ways to dynamically search for the best learning strategy as the number of tasks increase, implying some form of generalisation of the learning process to other problems.

@@

\acp{HH} implement a form of \index{meta-learning}meta-learning that is concerned with the selection of the best heuristic from a pool of heuristics to solve a given problem. In the context of population-based \acp{HH}, an entity pool exists that represent a pool of candidate solutions to the given problem. Each entity in the entity pool is assigned its own low-level \index{heuristic}heuristic from the heuristic pool.

@@

Figure \ref{fig:heuristics:hh:hh_algorithm} shows the separation of information that is accessible to the entity pool and heuristic pool. At a higher level abstraction, the high level heuristic implemented by \acp{HH}, do not make use of domain specific information such as an entity's position or gradient from the search space.

@@

\acp{HH} make use of feedback from the search process to adapt and guide the search process. Some \acp{HH} implement a learning mechanism that utilises this feedback. Learning \acp{HH} can implement a form of \index{online learning}\textit{online learning} or \index{offline learning}\textit{offline learning}.

@@

\acp{HH} that implement \index{offline learning}offline learning, implement a form of learning where knowledge, in the form of rules or programs, is gathered from a set of training instances and is then applied independently from the search process itself. Examples of offline learning approaches within hyper-heuristics include \textit{learning classifier systems} and \textit{case-based reasoning}.

@@

Selection \acp{HH} consist of two components, including a low-level \index{heuristic}heuristic selection strategy and a move acceptance strategy. Low-level \index{heuristic}heuristic selection can be done in a simple, \textit{non-adaptive} way. No learning is involved in these approaches. For non-adaptive techniques, heuristic selection is based on a predefined heuristic ordering that is generated either randomly or in a ordered cycle that is repeated throughout the optimisation process~\cite{ref:cowling:2000}. As an alternative, heuristic selection may incorporate an \textit{adaptive} (\index{online learning}\textit{online learning}) mechanism, based on the probabilistic weighting of the low-level heuristics~\cite{ref:burke:2003} or some type of performance statistics~\cite{ref:cowling:2000}.

A move acceptance strategy is the mechanism by which the application of a low-level \index{heuristic}heuristic is either accepted or rejected. In general, a move is accepted or rejected based on the quality of the move and the current solution during a single point search.~\citeauthor{ref:burke:2003}~\cite{ref:burke:2003} mention that many move acceptance strategies have been explored within \acp{HH}.

@@

This chapter provided the reader with the necessary background information on \index{meta-learning}meta-learning and \acp{HH}. Formal definitions were given, followed by detailed discussions. A modern classification scheme for \acs{HH} as proposed by~\citeauthor{ref:burke:2010}~\cite{ref:burke:2010} was presented in detail. Distinction is made between the source of feedback information used during learning and the nature of the heuristic search space. Detailed discussions on selection and generation heuristics are provided, followed by a further classification of construction and perturbation mechanisms.
